---
Title: Cleaning Raw Microclimate Data
Format: html
Authors: M.R.Verschoor
Date: 2024-03-14
editor_options: 
  chunk_output_type: console
---

## Set-up

Notes: ALWAYS TRY TO RUN BEFORE CODING SO THAT PACKAGES ARE THE RIGHT VERSION AND SYNCHED WITH THE RENV ENVIRONMENT

```{r renv environment}

# clear everything in the R environment
rm(list = ls())

# restore the library
renv::restore()

# Load configuration settings for collaborative data loading
source("config.R")  

```


## Loading and installing packages

```{r packages}

# Loading and installing the required Packages

#General
library(tidyverse)

library(data.table)

library(fasttime)

library(lubridate)

```

## Loading in Microclimate data

Purpose of chunk: Import and clean the microclimate data 

Notes: Here, all data .csv files are imported and stored in a single data frame. To allow identification of the individual loggers as the logger data is stored in folders corresponding to the name of the containers. The functions below will name the folder according to the file location. 

NOT NEEDED IF ALREADY MADE, CONINUE TO CHUNK BELOW (SAVES LOT OF TIME)

```{r Loading in Microclimate data}

# Loading Bioclive design
bioclive_design <- load_data("bioclive_design.csv") |>
  select(-V1) |>
  mutate(subplot = paste0(plot, subplot)) |>
  mutate(
    dist_wall = case_when(
      plot %in% 1:11 | plot %in% c(12, 22, 23, 33, 34, 44, 55, 66, 77, 45, 56, 67, 78:88) ~ 1,
      plot %in% 13:21 | plot %in% 68:76 | plot %in% c(24, 35, 46, 57, 32, 43, 54, 65) ~ 2,
      plot %in% 25:31 | plot %in% 58:64 | plot %in% c(36, 47, 53, 42) ~ 3,
      plot %in% 37:41 | plot %in% 48:52 ~ 4,
      TRUE ~ NA_real_ # If no condition matches, return NA
    )
  )

# Reading in the microclimate data

# Tomst Probe data cleaning for analysis
# reading in the data and binding the different files

# Use the data_path root and specify the sub directories dynamically
# Define the directory for Probes2023-2024 within data_path
probe_data_dir <- file.path(data_path, "Data/Raw/Probes/")

# List all .csv files in the specified directory and its sub directories
tms.sm.files <- list.files(probe_data_dir, pattern = "\\.csv$", recursive = TRUE, full.names = TRUE)

# Initialize an empty list to store data frames, setting length based on the number of files
tms.sm.list <- vector(mode = "list", length = length(tms.sm.files))

# Function to read and name combo using fread for faster reading
read_and_name_combo <- function(TMSpath){
  data <- data.table::fread(TMSpath, sep = ";", header = FALSE)
  data[, probe := basename(TMSpath)]
  data[, subplot := basename(dirname(TMSpath))]
  return(data)
}

# Loop over all files and read each one, storing it in tms.sm.list
for (i in seq_along(tms.sm.files)) {
  TMSpath <- tms.sm.files[[i]]
  tms.sm.list[[i]] <- read_and_name_combo(TMSpath)
}

# Bind all rows together into a single data.table
microclimate_shield <- data.table::rbindlist(tms.sm.list)

```

```{r Cleaning up the probe data}

# Cleaning up the probe data
# Lowercase plots and format the probe code
microclimate_shield[, subplot := tolower(subplot)]
microclimate_shield[, probe := substr(probe, 6, 13)]

# Rename columns using data.table's `setnames` for efficiency
data.table::setnames(microclimate_shield, old = names(microclimate_shield)[1:9],
         new = c("index", "DateTime", "Timezone", "T1", "T2", "T3", "Moisture", "Shake", "ErrFlag"))

microclimate_shield[, V10 := NULL]  # Remove the unnecessary column

# Convert columns T1, T2, and T3 to numeric in one go
microclimate_shield[, c("T1", "T2", "T3") := lapply(.SD, as.numeric), .SDcols = c("T1", "T2", "T3")]

# Convert DateTime to POSIXct with timezone "UTC" (adjust if different format)
microclimate_shield[, DateTime := fastPOSIXct(DateTime, tz = "UTC")]

# Add a 2-hour timezone difference
microclimate_shield[, DateTime := DateTime + hours(2)]

# Extract Date and Time from the adjusted DateTime
microclimate_shield[, Date := as.Date(DateTime)]
microclimate_shield[, Time := format(DateTime, "%H:%M:%S")]

# Extract Month and year from the adjusted DateTime and add it as a new column
microclimate_shield[, Month := format(DateTime, "%B %Y")]

# Remove the Timezone column and any extra columns
microclimate_shield[, Timezone := NULL] 

# getting the correct dates
microclimate_shield = microclimate_shield[microclimate_shield$DateTime >= "2024-08-05 00:00:00",]

# Rearrange columns
data.table::setcolorder(microclimate_shield, c("index", "DateTime", "Date", "Time", "Month", "probe", "subplot", "T1", "T2", "T3", "Moisture", "Shake", "ErrFlag"))

# Add shield presence/absence column based on `subplot`
microclimate_shield[, midshield := ifelse(
    substr(subplot, nchar(subplot), nchar(subplot)) %in% c("a", "c") | subplot == "65d", "present",
    ifelse(substr(subplot, nchar(subplot), nchar(subplot)) %in% c("b", "d") | subplot == "65c", "absent", NA_character_)
)]

```

## Processing Microclimate data and calculating the SWC

Purpose of chunk: Format Logger data and calibrate TMS raw moisture signal to volumetric Soil Water content (SWC, %).

Notes: Resulting SWC is expressed as a fraction between 0 and 1.

```{r Calculating SWC microclimate data}

# Universal TOMST calibration 
microclimate_shield <- microclimate_shield  |> 
  dplyr::mutate(SWC = round((Moisture^2 * -1.34e-08
 + Moisture * 0.0002496218 + -0.1578888
), 4))

# Checking weather max to determine the cut-off point for outliers 

# Read the CSV file
knmi_data <- read_csv("C:/Users/minkv/OneDrive - Universiteit Utrecht/Projects/2025_Verschoor_Limited impact of lower shield installation/Data/Raw/Daggegevens weer.txt")  %>%
  mutate(YYYYMMDD = as.Date(as.character(YYYYMMDD), format = "%Y%m%d")) |> # Convert YYYYMMDD to Date format
  filter(YYYYMMDD >= as.Date("2024-08-05") & YYYYMMDD < as.Date("2024-09-04")) |> 
  mutate(TX = TX / 10) |> 
  mutate(TN = TN / 10)

max(knmi_data$TX)
min(knmi_data$TN)

# Removing the night hours (21:00 - 06:00) from the data & removing decrepit measurements 
microclimate_shield_day <- microclimate_shield[
  Time >= "06:00" & Time <= "21:00" & 
  !(subplot %in% c("63B", "58A")) &
  (SWC >= 0.01 | is.na(SWC)) &
  (T1 <= 45 | is.na(T1)) &
  (T2 <= 45 | is.na(T2)) &
  (T3 <= 45 | is.na(T3)) &
  (T1 >= -20 | is.na(T1)) &
  (T2 >= -20 | is.na(T2)) &
  (T3 >= -20 | is.na(T3))
]

microclimate_shield$Time <- as.POSIXct(microclimate_shield$Time, format="%H:%M:%S")

# Assuming `Time` is a POSIXct or POSIXlt object, or it's in an easy-to-parse format
microclimate_shield_night <- microclimate_shield[
  (format(Time, "%H:%M:%S") >= "21:00:00" | format(Time, "%H:%M:%S") <= "06:00:00") & 
  !(subplot %in% c("63B", "58A")) &
  (SWC >= 0.01 | is.na(SWC)) &
  (T1 <= 45 | is.na(T1)) &
  (T2 <= 45 | is.na(T2)) &
  (T3 <= 45 | is.na(T3)) &
  (T1 >= -20 | is.na(T1)) &
  (T2 >= -20 | is.na(T2)) &
  (T3 >= -20 | is.na(T3))
]



```


```{r Creating the mean and max dataframe}

# Calculate the max per probe, the mean per probe for the total period

# Calculate the max per probe, the mean per probe for the total period, but group only by 'subplot' and 'Date'
microclimate_shield_total_day <- microclimate_shield_day %>%
  group_by(subplot) %>%  # Group by subplot and Date only
  dplyr::summarise(
    min_SWC = min(SWC, na.rm = TRUE),
    max_T1 = max(T1, na.rm = TRUE),
    max_T2 = max(T2, na.rm = TRUE),
    max_T3 = max(T3, na.rm = TRUE),
    min_SWC_1 = quantile(SWC, 0.01, na.rm = TRUE),
    max_T1_99 = quantile(T1, 0.99, na.rm = TRUE),
    max_T2_99 = quantile(T2, 0.99, na.rm = TRUE),
    max_T3_99 = quantile(T3, 0.99, na.rm = TRUE),
    min_SWC_5 = quantile(SWC, 0.05, na.rm = TRUE),
    max_T1_95 = quantile(T1, 0.95, na.rm = TRUE),
    max_T2_95 = quantile(T2, 0.95, na.rm = TRUE),
    max_T3_95 = quantile(T3, 0.95, na.rm = TRUE),
    min_SWC_10 = quantile(SWC, 0.10, na.rm = TRUE),
    max_T1_90 = quantile(T1, 0.90, na.rm = TRUE),
    max_T2_90 = quantile(T2, 0.90, na.rm = TRUE),
    max_T3_90 = quantile(T3, 0.90, na.rm = TRUE)
  ) %>%
  ungroup()  # Ungroup after summarisation

# Now join the original data back to keep 'probe' and 'midshield' columns
microclimate_shield_total_day <- microclimate_shield_total_day %>%
  left_join(
    microclimate_shield %>%
      select(subplot, probe, midshield) %>%
      distinct(),  # Keep distinct rows of 'probe' and 'midshield' for each subplot & Date
    by = c("subplot")
  ) |> 
  left_join(bioclive_design)

# Calculate the min per probe, the mean per probe for the total period, but group only by 'subplot' and 'Date'
microclimate_shield_total_night <- microclimate_shield_night %>%
  group_by(subplot) %>%
  dplyr::summarise(
    # absolute mins
    min_SWC = min(SWC, na.rm = TRUE),
    min_T1 = min(T1, na.rm = TRUE),
    min_T2 = min(T2, na.rm = TRUE),
    min_T3 = min(T3, na.rm = TRUE),

    # Lower quarantines (cold extremes)
    min_T1_01 = quantile(T1, 0.01, na.rm = TRUE),
    min_T2_01 = quantile(T2, 0.01, na.rm = TRUE),
    min_T3_01 = quantile(T3, 0.01, na.rm = TRUE),
    min_T1_05 = quantile(T1, 0.05, na.rm = TRUE),
    min_T2_05 = quantile(T2, 0.05, na.rm = TRUE),
    min_T3_05 = quantile(T3, 0.05, na.rm = TRUE),
    min_T1_10 = quantile(T1, 0.10, na.rm = TRUE),
    min_T2_10 = quantile(T2, 0.10, na.rm = TRUE),
    min_T3_10 = quantile(T3, 0.10, na.rm = TRUE),
    
   # SWC quantiles
    min_SWC_1 = quantile(SWC, 0.01, na.rm = TRUE),
    min_SWC_5 = quantile(SWC, 0.05, na.rm = TRUE),
    min_SWC_10 = quantile(SWC, 0.10, na.rm = TRUE)
  ) %>%
  ungroup()

# Now join the original data back to keep 'probe' and 'midshield' columns
microclimate_shield_total_night <- microclimate_shield_total_night %>%
  left_join(
    microclimate_shield %>%
      select(subplot, probe, midshield) %>%
      distinct(),  # Keep distinct rows of 'probe' and 'midshield' for each subplot & Date
    by = c("subplot")
  ) |> 
  left_join(bioclive_design)


```

```{r Creating the mean and max dataframe}

# Calculate the max per probe, the mean per probe per day

# Calculate the max per probe, the mean per probe per day, but group only by 'subplot' and 'Date'
microclimate_shield_date_day <- microclimate_shield_day %>%
  group_by(subplot, Date) %>%  # Group by subplot and Date only
   dplyr::summarise(
    min_SWC = min(SWC, na.rm = TRUE),
    max_T1 = max(T1, na.rm = TRUE),
    max_T2 = max(T2, na.rm = TRUE),
    max_T3 = max(T3, na.rm = TRUE),
    min_SWC_1 = quantile(SWC, 0.01, na.rm = TRUE),
    max_T1_99 = quantile(T1, 0.99, na.rm = TRUE),
    max_T2_99 = quantile(T2, 0.99, na.rm = TRUE),
    max_T3_99 = quantile(T3, 0.99, na.rm = TRUE),
    min_SWC_5 = quantile(SWC, 0.05, na.rm = TRUE),
    max_T1_95 = quantile(T1, 0.95, na.rm = TRUE),
    max_T2_95 = quantile(T2, 0.95, na.rm = TRUE),
    max_T3_95 = quantile(T3, 0.95, na.rm = TRUE),
    min_SWC_10 = quantile(SWC, 0.10, na.rm = TRUE),
    max_T1_90 = quantile(T1, 0.90, na.rm = TRUE),
    max_T2_90 = quantile(T2, 0.90, na.rm = TRUE),
    max_T3_90 = quantile(T3, 0.90, na.rm = TRUE)
  ) %>%
  ungroup()  # Ungroup after summarisation

# Now join the original data back to keep 'probe' and 'midshield' columns
microclimate_shield_date_day <- microclimate_shield_date_day %>%
  left_join(
    microclimate_shield %>%
      select(subplot, Date, probe, midshield) %>%
      distinct(),  # Keep distinct rows of 'probe' and 'midshield' for each subplot & Date
    by = c("subplot", "Date")
  ) |> 
  left_join(bioclive_design)

# Calculate the max per probe, the mean per probe for the total period, but group only by 'subplot' and 'Date'
microclimate_shield_date_night <- microclimate_shield_night %>%
  group_by(subplot, Date) %>%  # Group by subplot and Date only
  dplyr::summarise(
    # absolute mins
    min_SWC = min(SWC, na.rm = TRUE),
    min_T1 = min(T1, na.rm = TRUE),
    min_T2 = min(T2, na.rm = TRUE),
    min_T3 = min(T3, na.rm = TRUE),

    # Lower quarantines (cold extremes)
    min_T1_01 = quantile(T1, 0.01, na.rm = TRUE),
    min_T2_01 = quantile(T2, 0.01, na.rm = TRUE),
    min_T3_01 = quantile(T3, 0.01, na.rm = TRUE),
    min_T1_05 = quantile(T1, 0.05, na.rm = TRUE),
    min_T2_05 = quantile(T2, 0.05, na.rm = TRUE),
    min_T3_05 = quantile(T3, 0.05, na.rm = TRUE),
    min_T1_10 = quantile(T1, 0.10, na.rm = TRUE),
    min_T2_10 = quantile(T2, 0.10, na.rm = TRUE),
    min_T3_10 = quantile(T3, 0.10, na.rm = TRUE),
    
   # SWC quantiles
    min_SWC_1 = quantile(SWC, 0.01, na.rm = TRUE),
    min_SWC_5 = quantile(SWC, 0.05, na.rm = TRUE),
    min_SWC_10 = quantile(SWC, 0.10, na.rm = TRUE)
  ) %>%
  ungroup()

# Now join the original data back to keep 'probe' and 'midshield' columns
microclimate_shield_date_night <- microclimate_shield_date_night %>%
  left_join(
    microclimate_shield %>%
      select(subplot, probe, midshield) %>%
      distinct(),  # Keep distinct rows of 'probe' and 'midshield' for each subplot & Date
    by = c("subplot")
  ) |> 
  left_join(bioclive_design)


```

```{r Saving the data}

#Saving the cleaned data

# Define the output directory
output_dir <- file.path(data_path, "Data/Cleaned")

# Define the full file path for the CSV
output_file_1 <- file.path(output_dir, "microclimate_shield_total_day.csv")
output_file_2 <- file.path(output_dir, "microclimate_shield_total_night.csv")
output_file_3 <- file.path(output_dir, "microclimate_shield_date_night.csv")
output_file_4 <- file.path(output_dir, "microclimate_shield_date_day.csv")

# Save the filtered data to a CSV file using fwrite
data.table::fwrite(microclimate_shield_total_day, file = output_file_1)
data.table::fwrite(microclimate_shield_total_night, file = output_file_2)
data.table::fwrite(microclimate_shield_date_night, file = output_file_3)
data.table::fwrite(microclimate_shield_date_day, file = output_file_4)

# Confirmation message
cat("Data saved to:", output_file_1, "\n")
cat("Data saved to:", output_file_2, "\n")
cat("Data saved to:", output_file_3, "\n")
cat("Data saved to:", output_file_4, "\n")
```

