---
Title: Cleaning Raw Microclimate Data
Format: html
Authors: M.R.Verschoor
Date: 2024-03-14
editor_options: 
  chunk_output_type: console
---

## Set-up

Notes: ALWAYS TRY TO RUN BEFORE CODING SO THAT PACKAGES ARE THE RIGHT VERSION AND SYNCHED WITH THE RENV ENVIRONMENT

## Loading and installing packages

```{r renv environment}

# clear everything in the R environment
rm(list = ls()) # Optional: some prefer a clean slate, others find it risky
renv::restore() # restore the library
#renv::snapshot() # snapshot the library to update renv.lock if needed

## --- Load Packages ---
library(marginaleffects)
library(data.table)
library(broom.mixed)
library(officer)
library(flextable)
library(here)
library(fasttime)
library(lubridate)
library(tidyverse)
library(MetBrewer)
library(ggh4x)
library(ggsignif)
library(ggdist)
library(patchwork)
library(brms)
library(emmeans)
library(extrafont)
library(bayesplot)
library(posterior)


## --- Directory Setup ---
# Use here() so this works regardless of where the script is launched

data_dir <- here("data")
fig_dir  <- here("data", "figures", "revisions")

if (!dir.exists(data_dir)) dir.create(data_dir)
if (!dir.exists(fig_dir))  dir.create(fig_dir, recursive = TRUE)

```

## Loading in Microclimate data

Purpose of chunk: Import and clean the microclimate data 

Notes: Here, all data .csv files are imported and stored in a single data frame. To allow identification of the individual loggers as the logger data is stored in folders corresponding to the name of the containers. The functions below will name the folder according to the file location. 

NOT NEEDED IF ALREADY MADE, CONINUE TO CHUNK BELOW (SAVES LOT OF TIME)

```{r Loading in Data Zenodo}

# These files are hosted on Zenodo at DOI: 10.5281/zenodo.16677435
options(timeout = 600) # Increase timeout to 10 minutes

# Define URLs to files on Zenodo
microclimate_url <- "https://zenodo.org/record/16677435/files/Bioclive_microclimate_shield_2024.csv?download=1"
design_url <- "https://zenodo.org/record/16677435/files/bioclive_design.csv?download=1"
weather_url <- "https://zenodo.org/records/18482920/files/TheBilt%20Weather%20station%20data%20KNMI.txt?download=1"

# Define local filenames
microclimate_file <- "data/Bioclive_microclimate_shield_2024.csv"
design_file <- "data/bioclive_design.csv"
weather_file <- "data/TheBilt Weather station data KNMI.txt"

# Create data directory if it doesn't exist
if (!dir.exists("data")) dir.create("data")

# Download microclimate data if needed
if (!file.exists(microclimate_file)) {
  message("Downloading microclimate data...")
  download.file(microclimate_url, destfile = microclimate_file, mode = "wb")
} else {
  message("Microclimate data already exists locally.")
}

# Download design data if needed
if (!file.exists(design_file)) {
  message("Downloading design file...")
  download.file(design_url, destfile = design_file, mode = "wb")
} else {
  message("Design file already exists locally.")
}

# Download weather data if needed
if (!file.exists(weather_file)) {
  message("Downloading weather file...")
  download.file(weather_url, destfile = weather_file, mode = "wb")
} else {
  message("Weather file already exists locally.")
}

# Now load the data
microclimate_shield <- readr::read_csv(microclimate_file)
bioclive_design <- readr::read_csv(design_file)
weather_data <- readr::read_csv(weather_file)

```

```{r Adding wall distance to Microclimate data}

# Loading Bioclive design
bioclive_design <- bioclive_design|>
  select(-...1) |>
  mutate(subplot = paste0(plot, subplot)) |>
  mutate(
    dist_wall = case_when(
      plot %in% 1:11 | plot %in% c(12, 22, 23, 33, 34, 44, 55, 66, 77, 45, 56, 67, 78:88) ~ 1,
      plot %in% 13:21 | plot %in% 68:76 | plot %in% c(24, 35, 46, 57, 32, 43, 54, 65) ~ 2,
      plot %in% 25:31 | plot %in% 58:64 | plot %in% c(36, 47, 53, 42) ~ 3,
      plot %in% 37:41 | plot %in% 48:52 ~ 4,
      TRUE ~ NA_real_ # If no condition matches, return NA
    )
  )

```

```{r Cleaning up the probe data}

# Cleaning up the probe data
# Lowercase plots and format the probe code

# Convert to data.table
setDT(microclimate_shield)

microclimate_shield[, subplot := tolower(subplot)]
microclimate_shield[, probe := substr(probe, 6, 13)]

# Rename columns using data.table's `setnames` for efficiency
data.table::setnames(microclimate_shield, old = names(microclimate_shield)[1:9],
         new = c("index", "DateTime", "Timezone", "T1", "T2", "T3", "Moisture", "Shake", "ErrFlag"))

microclimate_shield[, V10 := NULL]  # Remove the unnecessary column

# Convert columns T1, T2, and T3 to numeric in one go
microclimate_shield[, c("T1", "T2", "T3") := lapply(.SD, as.numeric), .SDcols = c("T1", "T2", "T3")]

# Convert DateTime to POSIXct with timezone "UTC" (adjust if different format)
microclimate_shield[, DateTime := fastPOSIXct(DateTime, tz = "UTC")]

# Add a 2-hour timezone difference
microclimate_shield[, DateTime := DateTime + hours(2)]

# Extract Date and Time from the adjusted DateTime
microclimate_shield[, Date := as.Date(DateTime)]
microclimate_shield[, Time := format(DateTime, "%H:%M:%S")]

# Extract Month and year from the adjusted DateTime and add it as a new column
microclimate_shield[, Month := format(DateTime, "%B %Y")]

# Remove the Timezone column and any extra columns
microclimate_shield[, Timezone := NULL] 

# getting the correct dates
microclimate_shield = microclimate_shield[microclimate_shield$DateTime >= "2024-08-05 00:00:00",]

# Rearrange columns
data.table::setcolorder(microclimate_shield, c("index", "DateTime", "Date", "Time", "Month", "probe", "subplot", "T1", "T2", "T3", "Moisture", "Shake", "ErrFlag"))

# Add shield presence/absence column based on `subplot`
microclimate_shield[, midshield := ifelse(
    substr(subplot, nchar(subplot), nchar(subplot)) %in% c("a", "c") | subplot == "65d", "present",
    ifelse(substr(subplot, nchar(subplot), nchar(subplot)) %in% c("b", "d") | subplot == "65c", "absent", NA_character_)
)]

```

```{r Weather station data}
## The bilt weather station data

# Loading the bilt weather station data 
weather_data <- weather_data %>%
  # Select only the variables you need
  # TX: Max Temp, TG: Mean Temp, FHX: Max Wind, Q: Radiation, SQ: Sunshine
  select(YYYYMMDD, TX, TG, FHX, Q, SQ) %>%
  rename(
    Date = YYYYMMDD,
    max_temp = TX,             # Maximum temperature (in 0.1 °C)
    mean_temp = TG,            # Mean temperature (in 0.1 °C)
    max_windspeed = FHX,       # Highest hourly average windspeed (in 0.1 m/s)
    radiation = Q,             # Global radiation (in J/cm2)
    sunshine_min = SQ          # Sunshine duration (in 0.1 hours)
  ) %>%
  mutate(
    # 1. Convert Date string to Date object
    Date = as.Date(as.character(Date), format = "%Y%m%d"),
    
    # 2. Convert Temperatures from 0.1 °C to °C
    max_temp = max_temp * 0.1,
    mean_temp = mean_temp * 0.1,
    
    # 3. Convert Max Windspeed from 0.1 m/s to m/s
    max_windspeed = max_windspeed * 0.1,
    
    # 4. Convert Sunshine from 0.1 hours to Minutes
    # Formula: (Raw Value * 0.1) * 60  => Raw Value * 6
    sunshine_min = sunshine_min * 6,
    
    # 5. Radiation (Q) remains in J/cm2
    radiation = radiation 
  )%>%
  filter(Date >= as.Date("2024-08-05") & Date <= as.Date("2024-09-04"))

```

## Processing Microclimate data and calculating the SWC

Purpose of chunk: Format Logger data and calibrate TMS raw moisture signal to volumetric Soil Water content (SWC, %).

Notes: Resulting SWC is expressed as a fraction between 0 and 1.

```{r Calculating SWC microclimate data}

# Universal TOMST calibration 
microclimate_shield <- microclimate_shield  |> 
  dplyr::mutate(SWC = round((Moisture^2 * -1.34e-08
 + Moisture * 0.0002496218 + -0.1578888
), 4))

microclimate_shield$Time <- as.POSIXct(microclimate_shield$Time, format="%H:%M:%S")

# Filtering for Peak Day (6 hours) + Quality Control
microclimate_shield_day <- microclimate_shield[
  format(Time, "%H:%M:%S") >= "11:00:00" & 
  format(Time, "%H:%M:%S") <= "17:00:00" & 
  !(subplot %in% c("63B", "58A")) &
  (SWC > 0 | is.na(SWC)) &
  # Temperature bounds (using 60 as per your last snippet)
  (T1 <= 60 & T1 >= -20 | is.na(T1)) &
  (T2 <= 60 & T2 >= -20 | is.na(T2)) &
  (T3 <= 60 & T3 >= -20 | is.na(T3))
]

# Filtering for Deep Night (23:00 - 05:00)
# Filtering for Deep Night (6 hours) + Quality Control
# Filtering for Deep Night (23:00 - 05:00)
microclimate_shield_night <- microclimate_shield[
  (format(Time, "%H:%M:%S") >= "23:00:00" | format(Time, "%H:%M:%S") <= "05:00:00") & 
  !(subplot %in% c("63B", "58A")) &
  (SWC > 0 | is.na(SWC)) &
  (T1 <= 60 & T1 >= -20 | is.na(T1)) &
  (T2 <= 60 & T2 >= -20 | is.na(T2)) &
  (T3 <= 60 & T3 >= -20 | is.na(T3))
]

```

```{r Creating the mean and max dataframe}

# Calculate the max per probe, the mean per probe for the total period

# Calculate the max per probe, the mean per probe for the total period, but group only by 'subplot' and 'Date'
microclimate_shield_total_day <- microclimate_shield_day %>%
  group_by(subplot) %>%  # Group by subplot and Date only
  dplyr::summarise(
    min_SWC = min(SWC, na.rm = TRUE),
    max_T1 = max(T1, na.rm = TRUE),
    max_T2 = max(T2, na.rm = TRUE),
    max_T3 = max(T3, na.rm = TRUE),
    min_SWC_1 = quantile(SWC, 0.01, na.rm = TRUE),
    max_T1_99 = quantile(T1, 0.99, na.rm = TRUE),
    max_T2_99 = quantile(T2, 0.99, na.rm = TRUE),
    max_T3_99 = quantile(T3, 0.99, na.rm = TRUE),
    min_SWC_5 = quantile(SWC, 0.05, na.rm = TRUE),
    max_T1_95 = quantile(T1, 0.95, na.rm = TRUE),
    max_T2_95 = quantile(T2, 0.95, na.rm = TRUE),
    max_T3_95 = quantile(T3, 0.95, na.rm = TRUE),
    min_SWC_10 = quantile(SWC, 0.10, na.rm = TRUE),
    max_T1_90 = quantile(T1, 0.90, na.rm = TRUE),
    max_T2_90 = quantile(T2, 0.90, na.rm = TRUE),
    max_T3_90 = quantile(T3, 0.90, na.rm = TRUE)
  ) %>%
  ungroup()  # Ungroup after summarisation

# Now join the original data back to keep 'probe' and 'midshield' columns
microclimate_shield_total_day <- microclimate_shield_total_day %>%
  left_join(
    microclimate_shield %>%
      select(subplot, probe, midshield) %>%
      distinct(),  # Keep distinct rows of 'probe' and 'midshield' for each subplot & Date
    by = c("subplot")
  ) |> 
  left_join(bioclive_design)

# Calculate the min per probe, the mean per probe for the total period, but group only by 'subplot' and 'Date'
microclimate_shield_total_night <- microclimate_shield_night %>%
  group_by(subplot) %>%
  dplyr::summarise(
    # absolute mins
    min_SWC = min(SWC, na.rm = TRUE),
    min_T1 = min(T1, na.rm = TRUE),
    min_T2 = min(T2, na.rm = TRUE),
    min_T3 = min(T3, na.rm = TRUE),

    # Lower quarantines (cold extremes)
    min_T1_01 = quantile(T1, 0.01, na.rm = TRUE),
    min_T2_01 = quantile(T2, 0.01, na.rm = TRUE),
    min_T3_01 = quantile(T3, 0.01, na.rm = TRUE),
    min_T1_05 = quantile(T1, 0.05, na.rm = TRUE),
    min_T2_05 = quantile(T2, 0.05, na.rm = TRUE),
    min_T3_05 = quantile(T3, 0.05, na.rm = TRUE),
    min_T1_10 = quantile(T1, 0.10, na.rm = TRUE),
    min_T2_10 = quantile(T2, 0.10, na.rm = TRUE),
    min_T3_10 = quantile(T3, 0.10, na.rm = TRUE),
    
   # SWC quantiles
    min_SWC_1 = quantile(SWC, 0.01, na.rm = TRUE),
    min_SWC_5 = quantile(SWC, 0.05, na.rm = TRUE),
    min_SWC_10 = quantile(SWC, 0.10, na.rm = TRUE)
  ) %>%
  ungroup()

# Now join the original data back to keep 'probe' and 'midshield' columns
microclimate_shield_total_night <- microclimate_shield_total_night %>%
  left_join(
    microclimate_shield %>%
      select(subplot, probe, midshield) %>%
      distinct(),  # Keep distinct rows of 'probe' and 'midshield' for each subplot & Date
    by = c("subplot")
  ) |> 
  left_join(bioclive_design)


```

```{r Creating the mean and max dataframe}

# Calculate the max per probe, the mean per probe per day

# Calculate the max per probe, the mean per probe per day, but group only by 'subplot' and 'Date'
microclimate_shield_date_day <- microclimate_shield_day %>%
  group_by(subplot, Date) %>%  # Group by subplot and Date only
   dplyr::summarise(
    min_SWC = min(SWC, na.rm = TRUE),
    max_T1 = max(T1, na.rm = TRUE),
    max_T2 = max(T2, na.rm = TRUE),
    max_T3 = max(T3, na.rm = TRUE),
    min_SWC_1 = quantile(SWC, 0.01, na.rm = TRUE),
    max_T1_99 = quantile(T1, 0.99, na.rm = TRUE),
    max_T2_99 = quantile(T2, 0.99, na.rm = TRUE),
    max_T3_99 = quantile(T3, 0.99, na.rm = TRUE),
    min_SWC_5 = quantile(SWC, 0.05, na.rm = TRUE),
    max_T1_95 = quantile(T1, 0.95, na.rm = TRUE),
    max_T2_95 = quantile(T2, 0.95, na.rm = TRUE),
    max_T3_95 = quantile(T3, 0.95, na.rm = TRUE),
    min_SWC_10 = quantile(SWC, 0.10, na.rm = TRUE),
    max_T1_90 = quantile(T1, 0.90, na.rm = TRUE),
    max_T2_90 = quantile(T2, 0.90, na.rm = TRUE),
    max_T3_90 = quantile(T3, 0.90, na.rm = TRUE)
  ) %>%
  ungroup()  # Ungroup after summarisation

# Now join the original data back to keep 'probe' and 'midshield' columns
microclimate_shield_date_day <- microclimate_shield_date_day %>%
  left_join(
    microclimate_shield %>%
      select(subplot, Date, probe, midshield) %>%
      distinct(),  # Keep distinct rows of 'probe' and 'midshield' for each subplot & Date
    by = c("subplot", "Date")
  ) |> 
  left_join(bioclive_design)

# Calculate the max per probe, the mean per probe for the total period, but group only by 'subplot' and 'Date'
microclimate_shield_date_night <- microclimate_shield_night %>%
  group_by(subplot, Date) %>%  # Group by subplot and Date only
  dplyr::summarise(
    # absolute mins
    min_SWC = min(SWC, na.rm = TRUE),
    min_T1 = min(T1, na.rm = TRUE),
    min_T2 = min(T2, na.rm = TRUE),
    min_T3 = min(T3, na.rm = TRUE),

    # Lower quarantines (cold extremes)
    min_T1_01 = quantile(T1, 0.01, na.rm = TRUE),
    min_T2_01 = quantile(T2, 0.01, na.rm = TRUE),
    min_T3_01 = quantile(T3, 0.01, na.rm = TRUE),
    min_T1_05 = quantile(T1, 0.05, na.rm = TRUE),
    min_T2_05 = quantile(T2, 0.05, na.rm = TRUE),
    min_T3_05 = quantile(T3, 0.05, na.rm = TRUE),
    min_T1_10 = quantile(T1, 0.10, na.rm = TRUE),
    min_T2_10 = quantile(T2, 0.10, na.rm = TRUE),
    min_T3_10 = quantile(T3, 0.10, na.rm = TRUE),
    
   # SWC quantiles
    min_SWC_1 = quantile(SWC, 0.01, na.rm = TRUE),
    min_SWC_5 = quantile(SWC, 0.05, na.rm = TRUE),
    min_SWC_10 = quantile(SWC, 0.10, na.rm = TRUE)
  ) %>%
  ungroup()

# Now join the original data back to keep 'probe' and 'midshield' columns
microclimate_shield_date_night <- microclimate_shield_date_night %>%
  left_join(
    microclimate_shield %>%
      select(subplot, probe, midshield) %>%
      distinct(),  # Keep distinct rows of 'probe' and 'midshield' for each subplot & Date
    by = c("subplot")
  ) |> 
  left_join(bioclive_design)

# merging with weather data

microclimate_shield_date_day <- microclimate_shield_date_day |> 
  left_join(weather_data, by="Date")

microclimate_shield_date_night <- microclimate_shield_date_night |>
  left_join(weather_data, by="Date")



```

```{r Preparing the data for analysis}

# Ensure that Date is in Date format
microclimate_shield_date_day$Date <- as.Date(microclimate_shield_date_day$Date)
microclimate_shield_date_night$Date <- as.Date(microclimate_shield_date_night$Date)

microclimate_shield_date_day$yday <- format(microclimate_shield_date_day$Date, "%j")
microclimate_shield_date_day$yday = as.numeric(microclimate_shield_date_day$yday)

microclimate_shield_date_night$yday <- format(microclimate_shield_date_night$Date, "%j")
microclimate_shield_date_night$yday = as.numeric(microclimate_shield_date_night$yday)

microclimate_shield_date_day$midshield_binomial<-as.integer(if_else(microclimate_shield_date_day$midshield=="present", 1, 0))

microclimate_shield_date_night$midshield_binomial<-as.integer(if_else(microclimate_shield_date_night$midshield=="present", 1, 0))

microclimate_shield_total_day$midshield_binomial<-as.integer(if_else(microclimate_shield_total_day$midshield=="present", 1, 0))

microclimate_shield_total_night$midshield_binomial<-as.integer(if_else(microclimate_shield_total_night$midshield=="present", 1, 0))
```

## Setting up the color palette

```{r Setting up the color palette and fonts}

# Load the color palette and ensure it has enough colors for the unique levels in `treatment`
# Define the specific colors for each treatment
palette <- met.brewer("Lakota", n = 6, type = "discrete")
# Create the expanded palette for the gradient (12 colors for 12 species)

shield_colors <- c(
  "present" = palette[5],
  "absent" = palette[3])


loadfonts(device = "all")    # Loads the fonts

```

## Analysis

```{r Helper function}

## --- 1. Enhanced Sensitivity Analysis Function ---
run_sensitivity <- function(base_model, target_vars, data, period_label, ...) {
  model_results_list <- list()
  
  for (var in target_vars) {
    message(paste0("Running sensitivity analysis for: ", var, " (", period_label, ")"))
    
    # Dynamically update the formula while keeping random effects and interactions
    new_formula <- update.formula(formula(base_model), as.formula(paste(var, "~ .")))
    
    # Update the model with speed arguments (...)
    mod <- update(base_model, formula = new_formula, newdata = data, ...)
    
    # Save RDS for archival
    saveRDS(mod, file = here("data", paste0("m_", var, "_", period_label, ".rds")))
    
    # Extract fixed effects for the comparison plots
    res <- broom.mixed::tidy(mod, effects = "fixed", conf.int = TRUE) %>%
      mutate(model_label = var)
    
    model_results_list[[var]] <- res
  }
  return(bind_rows(model_results_list))
}

## --- 2. Master Helper Function ---
process_model_group <- function(base_model, target_vars, data, label, group_name, ...) {
  
  # --- A. Standard Density PP-Check ---
  pp_dens <- pp_check(base_model, ndraws = 100) + 
    labs(
      x = "Temperature (°C)", 
      y = "Probability Density",
      title = paste("Distribution Check:", group_name),
      subtitle = "Observed data (y) vs. 100 model simulations (y_rep)"
    ) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold"), plot.title = element_text(hjust = 0.5))
  
  ggsave(here("data", "figures", paste0("pp_check_", label, ".png")), pp_dens, width = 7, height = 5)

  # --- B. LOO-PIT QQ-Plot (Calibration Check) ---
  # Requires library(loo)
  pp_qq <- pp_check(base_model, type = "loo_pit_qq") +
    labs(
      title = paste("Calibration Check (LOO-PIT):", group_name),
      subtitle = "Assessing predictive distribution calibration (Uniformity)",
      x = "Theoretical Uniform Quantiles",
      y = "Sample Data Quantiles"
    ) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold"), plot.title = element_text(hjust = 0.5))
    
  ggsave(here("data", "figures", paste0("pp_qq_", label, ".png")), pp_qq, width = 7, height = 5)

  # --- C. Run Sensitivity Loop ---
  all_results <- run_sensitivity(base_model, target_vars, data, label, ...)
  
  # --- D. Extract Main Model Results (Fixed + Hyperparameters) ---
  # Including 'ran_pars' ensures we get sd(Intercept) for Date, Wall distance, etc.
  base_res_fixed <- broom.mixed::tidy(base_model, effects = "fixed", conf.int = TRUE) %>%
    mutate(model_label = "Base (95th)")
  
  # We return a list containing the full tidy results for tables AND the combined results for plots
  return(list(
    plot_data = bind_rows(all_results, base_res_fixed),
    full_table = broom.mixed::tidy(base_model, effects = c("fixed", "ran_pars"), conf.int = TRUE)
  ))
}

# Define a weakly informative prior for all fixed effects
priors <- prior(cauchy(0, 0.5), class = "b")

# Define a weakly informative prior for all fixed effects
priors <- prior(cauchy(0, 0.5), class = "b")

```

```{r Analysis}

# --------- SECTION 1: DAY TOTAL ----------#

m_day_95 <- brm(max_T2_95 ~ midshield_binomial + (1|div) + (1|treatment) + (1|dist_wall), 
      data = microclimate_shield_total_day, prior = priors,  cores = 4, chains = 4, iter = 4000, warmup = 1000)

day_total_results <- process_model_group(
  base_model   = m_day_95,
  target_vars  = c("max_T2_90", "max_T2_99", "max_T2"),
  data         = microclimate_shield_total_day,
  label        = "day_total",
  group_name   = "Daytime Max Temp", 
  # --- ADD THESE SPEED ARGUMENTS ---
  cores = 4,
  iter = 4000 
)

summary(m_day_95)

# --------- SECTION 2: NIGHT TOTAL ----------#
m_night_05 <- brm(min_T2_05 ~ midshield_binomial + (1|div) + (1|treatment) + (1|dist_wall),  data = microclimate_shield_total_night, prior = priors, cores = , chains = 4, iter = 4000, warmup = 1000)

night_total_results <- process_model_group(
  base_model   = m_night_05,
  target_vars  = c("min_T2_10", "min_T2_01", "min_T2"),
  data         = microclimate_shield_total_night,
  label        = "night_total",
  group_name   = "Nighttime Min Temp", 
  # --- ADD THESE SPEED ARGUMENTS ---
  cores = 4,
  iter = 4000 
)

summary(m_night_05)

# --------- SECTION 3: DAY DAILY (Interactions) ----------#

microclimate_shield_date_day$rad_scaled <- scale(microclimate_shield_date_day$radiation)
microclimate_shield_date_day$wind_scaled <- scale(microclimate_shield_date_day$max_windspeed)
microclimate_shield_date_day$Date <- as.factor(microclimate_shield_date_day$Date)

m_day_date_95 <- brm(max_T2_95 ~ midshield_binomial * (rad_scaled + wind_scaled) + (1|div) +  (1|treatment) + (1|dist_wall) + (1|Date), data = microclimate_shield_date_day, prior = priors, cores = 4, chains = 4, iter = 4000, warmup = 1000)

day_date_results <- process_model_group(
  base_model   = m_day_date_95,
  target_vars  = c("max_T2_90", "max_T2_99", "max_T2"),
  data         = microclimate_shield_date_day,
  label        = "day_date",
  group_name   = "Daily Day (Interactions)", 
  # --- ADD THESE SPEED ARGUMENTS ---
  cores = 4,
  iter = 4000 
)

summary(m_day_date_95)

# --------- SECTION 4: NIGHT DAILY ----------#

microclimate_shield_date_night$rad_scaled <- scale(microclimate_shield_date_night$radiation)
microclimate_shield_date_night$wind_scaled <- scale(microclimate_shield_date_night$max_windspeed)
microclimate_shield_date_night$Date <- as.factor(microclimate_shield_date_night$Date)

m_night_date_05 <- brm(min_T2_05 ~ midshield_binomial * wind_scaled + (1|div) +  (1|treatment) + (1|dist_wall) + (1|Date), data = microclimate_shield_date_night, prior = priors, cores = 4, chains = 4, iter = 4000, warmup = 1000)

night_date_results <- process_model_group(
  base_model   = m_night_date_05,
  target_vars  = c("min_T2_10", "min_T2_01", "min_T2"),
  data         = microclimate_shield_date_night,
  label        = "night_date",
  group_name   = "Daily Night (Interactions)", 
  # --- ADD THESE SPEED ARGUMENTS ---
  cores = 4,
  iter = 4000 
)

summary(m_night_date_05)


# Save the models as .rds files
saveRDS(m_day_95, "m_day_95.rds")
saveRDS(m_night_05, "m_night_05.rds")
saveRDS(m_day_date_95, "m_day_date_95.rds")
saveRDS(m_night_date_05, "m_night_date_05.rds")


```

## Supplementary info

```{r PP_Check plots & Comparison plots}

# Helper function to generate and save a labeled pp_check
save_labeled_pp <- function(model, title, filename) {
  p <- pp_check(model, ndraws = 100) + 
    labs(
      title = paste("PP-Check:", title),
      subtitle = "Observed (y) vs. Simulated (y_rep)",
      x = "Temperature (°C)", 
      y = "Probability Density"
    ) +
    theme_bw() +
    theme(axis.title = element_text(face = "bold"), plot.title = element_text(hjust = 0.5))
  
  ggsave(here("data", "figures", filename), p, width = 6, height = 4, dpi = 300)
}

# USE THIS VERSION to compare 90th, 95th, 99th on one plot
save_robustness_comp <- function(model_group_results, title, filename) {
  
  # 1. Define the specific orders you requested
  day_order <- c("Base (95th)", "max_T2_90", "max_T2_99", "max_T2")
  night_order <- c("Base (05th)", "min_T2_10", "min_T2_01", "min_T2")
  
  # 2. Check if this is a Night model (contains "min") 
  is_night <- any(grepl("min", model_group_results$model_label))
  
  plot_data <- model_group_results %>%
    # Fix the label for Nighttime Base so it's not "95th"
    mutate(model_label = if_else(is_night & model_label == "Base (95th)", "Base (05th)", model_label)) %>%
    # Set the factor levels based on the model type
    mutate(model_label = factor(model_label, levels = if(is_night) night_order else day_order))

  p <- ggplot(plot_data, aes(x = estimate, y = term, color = model_label)) +
    geom_vline(xintercept = 0, color = "black", linetype = "dashed", alpha = 0.5) +
    geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), 
                   height = 0.4, 
                   position = position_dodge(width = 0.6)) +
    geom_point(position = position_dodge(width = 0.6), size = 2.5) +
    labs(
      title = title,
      subtitle = "Sensitivity Analysis: Comparing Thresholds",
      x = "Effect Size (Temperature Difference °C)",
      y = "Parameter",
      color = "Model / Threshold"
    ) +
    theme_bw() +
    scale_color_brewer(palette = "Set1") + 
    theme(legend.position = "bottom", 
          axis.title = element_text(face = "bold"),
          plot.title = element_text(size = 14))

  ggsave(here("data", "figures", filename), p, width = 8, height = 6, dpi = 300)
}

# --- Execute Plot Generation ---

# --- EXECUTE PLOT GENERATION ---

# 1. Daytime Total Analysis
save_labeled_pp(m_day_95, "Daytime Max", "pp_check_day_total.png")
# Fix: Access the $plot_data element from the list
save_robustness_comp(day_total_results$plot_data, "Daytime Max Totals", "comparison_day_total.png")

# 2. Nighttime Total Analysis
save_labeled_pp(m_night_05, "Nighttime Min", "pp_check_night_total.png")
# Fix: Access the $plot_data element from the list
save_robustness_comp(night_total_results$plot_data, "Nighttime Min Totals", "comparison_night_total.png")

# 3. Daily Day Analysis (Interactions)
save_labeled_pp(m_day_date_95, "Daily Day (Interactions)", "pp_check_day_date.png")
# Fix: Access the $plot_data element from the list
save_robustness_comp(day_date_results$plot_data, "Daily Day Interactions", "comparison_day_date.png")

# 4. Daily Night Analysis (Interactions)
save_labeled_pp(m_night_date_05, "Daily Night (Interactions)", "pp_check_night_date.png")
# Fix: Access the $plot_data element from the list
save_robustness_comp(night_date_results$plot_data, "Daily Night Interactions", "comparison_night_date.png")

```

```{r Supplementary info}

# ----- Bayesian Results Table (Includes Fixed Effects & Hyperparameters) -----
make_bayesian_table <- function(model_data, title, subtitle = "Main Model") {
  
  # Handle list or model input
  if(is.list(model_data) && "full_table" %in% names(model_data)) {
    tab_data <- model_data$full_table
  } else {
    tab_data <- broom.mixed::tidy(model_data, effects = c("fixed", "ran_pars"), conf.int = TRUE)
  }

  tab <- tab_data %>%
    # Use the 'group' column to give the SDs unique names
    mutate(Parameter = case_when(
      effect == "ran_pars" & term == "sd__(Intercept)" ~ paste0("SD (Intercept: ", group, ")"),
      effect == "ran_pars" & term == "sd__Observation" ~ "SD (Residual)",
      # Clean up Intercept for fixed effects
      term == "(Intercept)" ~ "Intercept (Grand Mean)",
      TRUE ~ term
    )) %>%
    mutate(across(where(is.numeric), ~ round(.x, 3))) %>%
    select(Parameter, estimate, std.error, conf.low, conf.high) %>%
    rename(
      `Posterior Mean` = estimate,
      `Est. Error` = std.error,
      `95% Lower CI` = conf.low,
      `95% Upper CI` = conf.high
    )

  ft <- flextable::flextable(tab) %>%
    flextable::theme_vanilla() %>%
    flextable::set_caption(caption = paste(title, "-", subtitle)) %>%
    # Subtle gray background for the SD rows to separate them from fixed effects
    flextable::bg(i = ~ grepl("SD \\(", Parameter), bg = "gray97") %>%
    flextable::italic(i = ~ grepl("SD \\(", Parameter)) %>%
    flextable::autofit()
  
  return(ft)
}

# --- Build the master list for the Word Doc ---
# Note: 'main' is now the output from our new process_model_group
# --- Build the master list for the Word Doc ---
full_supplement_list <- list(
  
  "Daytime Maximum Temperature (Total Period)" = list(
    results   = day_total_results, 
    pp_path   = here("data", "figures", "pp_check_day_total.png"),
    qq_path   = here("data", "figures", "pp_qq_day_total.png"),
    comp_path = here("data", "figures", "comparison_day_total.png")
  ),

  "Nighttime Minimum Temperature (Total Period)" = list(
    results   = night_total_results,
    pp_path   = here("data", "figures", "pp_check_night_total.png"),
    qq_path   = here("data", "figures", "pp_qq_night_total.png"),
    comp_path = here("data", "figures", "comparison_night_total.png")
  ),

  "Daytime Maximum Temperature (Daily Interactions)" = list(
    results   = day_date_results, 
    pp_path   = here("data", "figures", "pp_check_day_date.png"),
    qq_path   = here("data", "figures", "pp_qq_day_date.png"),
    comp_path = here("data", "figures", "comparison_day_date.png")
  ),

  "Nighttime Minimum Temperature (Daily Interactions)" = list(
    results   = night_date_results,
    pp_path   = here("data", "figures", "pp_check_night_date.png"),
    qq_path   = here("data", "figures", "pp_qq_night_date.png"),
    comp_path = here("data", "figures", "comparison_night_date.png")
  ) # No comma here because it is the last item in the master list
)

# --- Generate the Document ---
doc <- read_docx()

iwalk(full_supplement_list, function(content, title) {
  
  doc <<- doc %>%
    body_add_par(title, style = "heading 1") %>%
    body_add_par("Main Bayesian Model Summary (Fixed & Random Effects)", style = "heading 2") %>%
    # FIX: Pass 'content$results' (the list), NOT 'content$results$full_table'
    body_add_flextable(make_bayesian_table(content$results, title))
  
  # A. Standard PP-Check
  if(!is.null(content$pp_path) && file.exists(content$pp_path)) {
    doc <<- doc %>% 
      body_add_par("Posterior Predictive Check: Distribution", style = "heading 2") %>%
      body_add_img(src = content$pp_path, width = 6, height = 4) %>%
      body_add_par("Figure Note: The dark line (y) is observed data; light lines (y_rep) are simulations. This shows how well the model captures the overall shape of the temperature distribution.", style = "Image Caption")
  }

  # B. LOO-PIT QQ-Plot (Calibration Check)
  if(!is.null(content$qq_path) && file.exists(content$qq_path)) {
    doc <<- doc %>% 
      body_add_par("Model Calibration: LOO-PIT QQ-Plot", style = "heading 2") %>%
      body_add_img(src = content$qq_path, width = 6, height = 4) %>%
      body_add_par("Figure Note: This plot assesses calibration. If the points follow the diagonal line, the model's uncertainty and probability of extremes are correctly estimated.", style = "Image Caption")
  }

  # C. Comparison Plot (Robustness)
  if(!is.null(content$comp_path) && file.exists(content$comp_path)) {
    doc <<- doc %>% 
      body_add_par("Sensitivity Analysis (Threshold Comparison)", style = "heading 2") %>%
      body_add_img(src = content$comp_path, width = 6, height = 4) %>%
      body_add_par("Figure Note: Comparison of effect sizes across 90th, 95th, and 99th percentiles. Consistency across thresholds demonstrates result robustness.", style = "Image Caption")
  }
  
  doc <<- doc %>% body_add_break()
})

print(doc, target = here("Manuscript", "Full_Automated_Supplement.docx"))
```

## Plots

```{r Plots}


# ------- Plot A: Daytime Violin Plot (no x-title) -------- # 

Plot_A_refined <- ggplot(microclimate_shield_total_day, aes(x = factor(midshield), y = max_T2_95, fill = midshield)) +
  # Use a half-violin (raincloud plot) to make it less crowded
  ggdist::stat_halfeye(adjust = .5, width = .3, .width = 0, justification = -.3, point_colour = NA) + 
  geom_boxplot(width = .1, outlier.shape = NA, alpha = 0.5) +
  # Add a horizontal line at the Intercept from your Bayesian model
  geom_hline(yintercept = 29.35, linetype = "dashed", alpha = 0.5) + 
  labs(subtitle = "A: Daytime (11:00 - 17:00)", y = "Max temp (°C)") +
   theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(family = "Arial", size = 16),
    axis.text.y = element_text(family = "Arial", size = 16),
    axis.title.y = element_text(family = "Arial", size = 16, margin = margin(r = 10)),
    legend.position = "none"
  ) +
  scale_color_manual(values = shield_colors) +
  scale_fill_manual(values = shield_colors)

Plot_A_refined

Plot_A <- ggplot(microclimate_shield_day, aes(x = factor(midshield, levels = c('present', 'absent')), y = max_T2_95, color = midshield, fill = midshield)) +
  geom_violin(position = position_dodge(width = 0.7), alpha = 0.3, width = 0.6) +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 0.5, position = position_dodge(width = 0.7), alpha = 0.3) +
  stat_summary(fun = mean, geom = "point", size = 4, shape = 16, position = position_dodge(width = 0.7)) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.3, position = position_dodge(width = 0.7)) +
  labs(y = "Max temp (°C) at 1 cm (06:00 - 21:00)") +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(family = "Arial", size = 16),
    axis.text.y = element_text(family = "Arial", size = 16),
    axis.title.y = element_text(family = "Arial", size = 16, margin = margin(r = 10)),
    legend.position = "none"
  ) +
  scale_color_manual(values = shield_colors) +
  scale_fill_manual(values = shield_colors)

# ------- Plot B: Daytime Time Series (no x-title) -------- # 
Plot_B <- ggplot(microclimate_shield_date_day, aes(x = Date, y = max_T2, color = midshield, fill = midshield)) +
  stat_summary(fun = mean, geom = "point", size = 4, shape = 16) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.3) +
  stat_summary(fun = mean, geom = "line", size = 1, alpha = 0.7, aes(group = midshield)) +
  labs(y = "Max daily temp (°C) at 1 cm (06:00 - 21:00)") +
  theme_minimal() +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, family = "Arial", size = 16),
    axis.text.y = element_text(family = "Arial", size = 16),
    axis.title.y = element_text(family = "Arial", size = 16, margin = margin(r = 10)),
    legend.position = "none"
  ) +
  scale_x_date(breaks = scales::date_breaks("3 days"),
               labels = scales::date_format("%b %d")) +
  scale_color_manual(values = shield_colors) +
  scale_fill_manual(values = shield_colors)

# ------- Plot C: Nighttime Violin Plot -------- # 
Plot_C <- ggplot(microclimate_shield_night, aes(x = factor(midshield, levels = c('present', 'absent')), y = min_T2_05, color = midshield, fill = midshield)) +
  geom_violin(position = position_dodge(width = 0.7), alpha = 0.3, width = 0.6) +
  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 0.5, position = position_dodge(width = 0.7), alpha = 0.3) +
  stat_summary(fun = mean, geom = "point", size = 4, shape = 16, position = position_dodge(width = 0.7)) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.3, position = position_dodge(width = 0.7)) +
  labs(x = "Shield configuration", y = "Min temp (°C) at 1 cm (21:00 - 06:00)") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(family = "Arial", size = 16),
    axis.title.x = element_text(family = "Arial", size = 16, margin = margin(t = 10)),
    axis.text.y = element_text(family = "Arial", size = 16),
    axis.title.y = element_text(family = "Arial", size = 16, margin = margin(r = 10)),
    legend.position = "none"
  ) +
  scale_color_manual(values = shield_colors) +
  scale_fill_manual(values = shield_colors)

# ------- Plot D: Nighttime Time Series -------- # 
Plot_D <- ggplot(microclimate_shield_date_night, aes(x = Date, y = min_T2_05, color = midshield, fill = midshield)) +
  stat_summary(fun = mean, geom = "point", size = 4, shape = 16) +
  stat_summary(fun.data = mean_cl_normal, geom = "errorbar", width = 0.3) +
  stat_summary(fun = mean, geom = "line", size = 1, alpha = 0.7, aes(group = midshield)) +
  labs(x = "Date", y = "Min daily temp (°C) at 1 cm (21:00 - 06:00)") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, family = "Arial", size = 16),
    axis.title.x = element_text(family = "Arial", size = 16, margin = margin(t = 10)),
    axis.text.y = element_text(family = "Arial", size = 16),
    axis.title.y = element_text(family = "Arial", size = 16, margin = margin(r = 10)),
    legend.position = "none"
  ) +
  scale_x_date(breaks = scales::date_breaks("3 days"),
               labels = scales::date_format("%b %d")) +
  scale_color_manual(values = shield_colors) +
  scale_fill_manual(values = shield_colors)

# ------- Combine and Adjust Layout -------- #
combined_plot <- (Plot_A + Plot_B) / (Plot_C + Plot_D) +
  theme(plot.tag = element_text(size = 20, face = "bold"),
        plot.tag.position = c(0, 1.1))  # More space between tag and plot

combined_plot

# Save to file
ggsave(file.path(data_path, "Plot_final.pdf"), combined_plot, width = 14, height = 12, dpi = 300, device = cairo_pdf)


```
